quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-05-15 paper-quollr.Rmd
Add algorithm section

@@

The paper is organized as follows. The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualization capabilities. In the application section, we illustrate the algorithm's functionality for studying clustering data structures. Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.

@@

@@

`scurve` is a $7\text{-}D$ simulated dataset. It is constructed by simulating $5000$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$. The remaining variables $X_4, X_5, X_6, X_7$ are all uniform error, with small variance. `scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors` is $46$ and `min_dist` is $0.9$. Each data set contains a unique ID column that maps `scurve` and `scurve_umap`.

@@

#### Scaling the Data

>>>>

Objective: From this, we can understand how an NLDR warps the high-dimensional space and fits the data.

@@

The paper is organized as follows. The next section introduces the implementation of the `quollr` package on CRAN, including a demonstration of the package's key functions and visualization capabilities. In the application section, we illustrate the algorithm's functionality for studying a clustering data structure. Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.

@@

Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, (6) identifying and removing low-density hexagons, and (7) lifting the model into higher dimensions. A detailed description of the algorithm can be found in **cite the methodology paper**.

Add a Figure explaining the steps

@@

`scurve` is a $7\text{-}D$ simulated dataset. It is constructed by simulating $5000$ observations from $\theta \sim U(-3\pi/2, 3\pi/2)$, $X_1 = \sin(\theta)$, $X_2 \sim U(0, 2)$ (adding thickness to the S), $X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)$. The remaining variables $X_4$, $X_5$, $X_6$, $X_7$ are all uniform error, with small variance. `scurve_umap` is the UMAP $2\text{-}D$ embedding for `scurve` data with `n_neighbors` is $46$ and `min_dist` is $0.9$. Each data set contains a unique ID column that maps `scurve` and `scurve_umap`.

@@

#### Scaling the data

### 2025-05-15 paper-quollr.tex
Add algorithm section

The paper is organized as follows. The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualization capabilities. In the application section, we illustrate the algorithm's functionality for studying clustering data structures. Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.

@@

@@

\texttt{scurve} is a \(7\text{-}D\) simulated dataset. It is constructed by simulating \(5000\) observations from \(\theta \sim U(-3\pi/2, 3\pi/2)\), \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\) (adding thickness to the S), \(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\). The remaining variables \(X_4, X_5, X_6, X_7\) are all uniform error, with small variance. \texttt{scurve\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{scurve} data with \texttt{n\_neighbors} is \(46\) and \texttt{min\_dist} is \(0.9\). Each data set contains a unique ID column that maps \texttt{scurve} and \texttt{scurve\_umap}.

@@

\paragraph{Scaling the Data}\label{scaling-the-data}

>>>>

The paper is organized as follows. The next section introduces the implementation of the \texttt{quollr} package on CRAN, including a demonstration of the package's key functions and visualization capabilities. In the application section, we illustrate the algorithm's functionality for studying a clustering data structure. Finally, we conclude the paper with a brief summary and discuss potential opportunities for using our algorithm.

@@

Our algorithm includes the following steps: (1) scaling the NLDR data, (2) computing configurations of a hexagon grid, (3) binning the data, (4) obtaining the centroids of each bin, (5) indicating neighboring bins with line segments that connect the centroids, (6) identifying and removing low-density hexagons, and (7) lifting the model into higher dimensions. A detailed description of the algorithm can be found in \textbf{cite the methodology paper}.

Add a Figure explaining the steps

@@

\texttt{scurve} is a \(7\text{-}D\) simulated dataset. It is constructed by simulating \(5000\) observations from \(\theta \sim U(-3\pi/2, 3\pi/2)\), \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\) (adding thickness to the S), \(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\). The remaining variables \(X_4\), \(X_5\), \(X_6\), \(X_7\) are all uniform error, with small variance. \texttt{scurve\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{scurve} data with \texttt{n\_neighbors} is \(46\) and \texttt{min\_dist} is \(0.9\). Each data set contains a unique ID column that maps \texttt{scurve} and \texttt{scurve\_umap}.

@@

\paragraph{Scaling the data}\label{scaling-the-data}
